\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-12-30 forwards-and-backwards.tex
AME edits

\KEYWORDS{population model ; interacting superprocess ; lookdown construction ; porous medium equation ; reaction-diffusion equation ; travelling waves ; genealogies ; Fisher-KPP equation}
\AMSSUBJ{60J25 ; 92D10 ; 92D15 ; 92D25 ; 92D40}
\AMSSUBJSECONDARY{60F05 ; 60G09 ; 60G55 ; 60G57 ; 60H15 ; 60J68}

@@

The history of a natural population is often only accessible indirectly, 
through patterns of genetic diversity:
from genetic data, one can try to infer the genealogical trees that describe how
individuals in a sample from the population are related, and these have been shaped 
by its history (see e.g., \citep{neigel1993application,kelleher2019inferring}).
It is therefore of interest to establish information about the distribution of genealogical trees 
under our population model, which we do with a lookdown construction. 
Lookdown constructions were first introduced in~\cite{donnelly/kurtz:1996} to provide 
a mechanism for retaining information about genealogical relationships between individuals
sampled from a population evolving according to the Moran model when passing to the 
infinite population limit.
Since then, they have been extended to a wide range of models. Of particular
relevance to our work here are the papers of~\cite{kurtz/rodrigues:2011} 
and~\cite{etheridge/kurtz:2019}, in which
lookdown constructions are provided for a wide variety of population models, including
spatially structured branching processes.

@@

In what follows we first study various scaling limits of the spatial population process,
and then turn our attention to lineages traced back through these populations.
First, in Section~\ref{sec: Model and main results},
we describe the model and the main results,
Theorems~\ref{thm:nonlocal_convergence}, \ref{thm:local_convergence}, and~\ref{thm:lineages}.
Next, in Section~\ref{sec:applications}, we discuss a few striking consequences of these results
regarding the behavior of genealogies in traveling waves,
the appearance of periodic ``clumps'' in seemingly homogeneous population models,
and identifiability of the underlying dynamics from a stationary population profile.
In Section~\ref{sec:heuristics}, we provide heuristic explanations
of why the theorems ought to be true,
and some key ideas behind them,
and in Section~\ref{sec:lookdown} we define and discuss the lookdown construction.
Proofs of the results begin in Section~\ref{sec:proofs},
which proves results for population models with nonlocal interactions,
while Section~\ref{subsec:one step convergence proof} gives the more difficult proof
for the case when interaction distances also go to zero in the limit.
Finally, Section~\ref{sec:lookdown_proofs} gives proofs for convergence of the lookdown process
and the associated results for the motion of lineages.
The Appendix contains a few more technical and less central lemmas.
The results are illustrated in a few places with individual-based simulations,
made using SLiM \citep{haller_slim_2019},
but these are provided for visualization and we do not embark on numerical study.

@@

\paragraph{Scaling limits with nonlocal interactions:}
Recall that the process $(\eta_t^N)_{t\geq 0}$ takes its values in the space
${\mathcal D}_{[0,\infty)}(\measures)$ of c\`adl\`ag
paths on $\measures$. We endow $\measures$ with the topology of weak convergence
and 
${\mathcal D}_{[0,\infty)}(\measures)$ with the Skorohod topology.
A sequence of processes taking values in 
${\mathcal D}_{[0,\infty)}(\measures)$ is said to be tight if the corresponding
sequence of distributions is tight,
i.e., that any infinite subsequence has a weakly convergent subsubsequence.
Our first main result establishes tightness of our rescaled population processes in 
the case in which interactions remain nonlocal under the scaling, and 
characterises limit points as solutions to a martingale problem.

@@

Now that we have established what we can say about how 
population density changes with time,
we turn to results on ancestral lineages,
i.e., how genealogical ancestry can be traced back across the landscape.
Informally,
a \emph{lineage} $(L_t^N)_{t \ge 0}$,
begun at a spatial location $L_0^N = x$
where there is a focal individual in the present day,
can be obtained for each time $t$ by
setting $L_t^N$ to be the spatial location of the individual alive at time $t$
before the present from
whom the focal individual is descended.
Since in our model individuals have only one parent, this is unambiguous.
Although we did not explicitly retain such information,
it is clear that for finite $N$, since individuals are born one at a time, 
one could construct the lineage $(L_t^N)_{t=0}^T$
given the history of the population $(\eta^N_t)_{t = 0}^T$,
for each starting location to which $\eta^N_T$ assigns positive mass.
It is less clear, however, how to formally retain such information when we 
pass to the scaling limit.

However, the \emph{lookdown construction} in Section~\ref{sec:lookdown}
does just this -- \llabel{whatis_lineage}
the construction enables us to recover information about ancestry
in the infinite population limit,
and thus gives a concrete meaning to $(L_t)_{t \ge 0}$. \llabel{Linf_mention}
Roughly speaking,
each particle is assigned a unique ``level'' from $[0,\infty)$
that functions as a label and thus allows reconstruction of lineages.
The key to the approach is that levels are assigned in such a way as to be exchangeable, 
so that sampling a finite number, $k$ say, of
individuals from a given 
region is equivalent to looking at the individuals in that region with the $k$ lowest levels.
Moreover, as we pass to the infinite population limit,
the collection of (individual, level) pairs converges,
as we show in Theorem~\ref{thm:lookdown_convergence}.
See~\citet{etheridge/kurtz:2019} for an introduction to these ideas.
In particular, even in the infinite
population limit, we can sample an individual
from a region (it will be the individual in that region with the lowest level) 
and trace its line of descent. 
This will allow us to calculate, for each $x$ and $y\in\IR^d$, the proportion of
the population at location $x$ in the present day population that are descended from 
a parent who was at location $y$ at time $t$ in the past. To make
sense of this in our framework, in Section~\ref{sec:lineages_proof}, 
we justify a weak reformulation of this idea.

@@

Ancestry in traveling waves for populations described by reaction-diffusion equations
has been studied before by various authors, \revpoint{1}{16}
although most work assumes the diffusion term is linear
(in our notation, $r$ and $\gamma$ are constant,
but $F$ can depend on density).
For instance, following \citet{hallatschek/nelson:2008},
various authors (e.g., \citet{roques2012allee,birzu/hallatschek/korolev:2017})
describe genetic diversity
and ancestry in (possibly noisy) traveling waves for reaction-diffusion equations
in a situation that covers our first two examples below.
However, we are not aware of previous work covering the third case 
with nonlinear diffusive term (the porous medium equation).

@@

It is interesting to compare the stationary distribution we have obtained here
to the expression that we'd get by setting $\meanq = -\wavespeed$
and using Corollary~\ref{cor:stationary_dist},
i.e., by giving each offspring a mean displacement that offsets the motion of the wave
(The model is different because this motion only affects offspring, not adults.)
In the Fisher-KPP and Allen-Cahn cases above, we get the same expressions,
but this is only the case because $r\equiv\gamma\equiv1$ in both.
For the PME with $\meanq = 1$ we have $\DG^* = \Delta + \grad$ 
and so the equation solved by the population density is
$$
    \partial_t \varphi = \Delta \varphi + 2 \varphi \grad \varphi + \varphi (1 - \varphi) ,
$$
which has a traveling wave solution of the same shape but moving at half the speed,
$\varphi(x,t) = w^{P}(x-t/2)$,
and a stationary distribution of the lineage relative to the wavefront of
$\pi(\xi) \propto e^{3\xi/2} (1 - e^{\xi/2})^3$.
% \comment{Details to back that last statement up, which we probably want to comment out?}
% (We can see that $w^P(x-t/2)$ is a traveling wave solution by using that
% $\partial_x w^P = -(1 - w^P)/2$;
% and then the generator of a linage in the moving frame is
% $\Lgen_{P'} = \varphi \partial_x^2 + (4 \partial_x \varphi + 1/2 + \varphi) \partial_x$,
% which has speed measure proportional to
% $\exp(\int^x 3(1/2 dx + d \log \varphi)) / \varphi$.)
% \begin{align*}
% m_{P'}(\xi)
%     &\propto \frac{1}{2(1-e^{\xi/2})} \exp\left(
%         \frac{3}{2} \int^\xi \left\{
%             1 - \frac{e^{x/2}}{1 - e^{x/2}}
%         \right\}dx
%     \right) .
% \end{align*}

@@

It is natural for applications to wonder about identifiability:
when can the observed quantities like population density
or certain summaries of lineage movement uniquely determine
the underlying demographic parameters?
Consider a deterministic,
continuous population generated by parameters $\gamma$, $r$, and $F$,
with $\meanq = 0$ and $\covq = I$.
Suppose it has a stationary profile $w(x)$, that must satisfy
$$
   r \Delta(\gamma w) + F w = 0 .
$$
It is easy to see that $w$ does not uniquely specify $\gamma$, $F$, and $r$:
let $\lambda(x)$ be a smooth, nonnegative function on $\IR^d$,
and let $\widetilde{r}(x, m) = \lambda(x) r(x, m)$ and $\widetilde{F}(x, m) = \lambda(x) F(x, m)$
(and, let $\widetilde{\gamma} = \gamma$).
Since $\mu = r \gamma - F/\theta$,
this corresponds to multiplying both establishment probabilities and death rates by $\lambda$.
Then the population with parameters $\widetilde{\gamma}$, $\widetilde{r}$, and $\widetilde{F}$
has the same stationary profile(s) as the original population.

@@

In this section we perform some preliminary calculations and use them 
to provide heuristic arguments for our main results,
to build intuition before the formal proofs.

@@

\paragraph{One-step convergence:} \label{sec:one_step_heuristics} \llabel{one_step_heuristics}
In order to pass directly to a classical PDE limit
in Theorem~\ref{thm:local_convergence}
we impose the stronger condition
that $\theta/N\epsilon^d\to 0$ and also require that
$\theta\epsilon^2\to\infty$. 
Recall that in this case, we take $\rho_F^\epsilon$ 
to be a symmetric Gaussian density with variance 
$\epsilon^2$. The condition $\theta\epsilon^2\to\infty$ 
ensures that $\epsilon^2$ is large enough relative to $1/\theta$
that the regularity gained by smoothing our population density by convolution with
$\rho_\epsilon$ is preserved under the dynamics dictated by $q_{\theta}$.
To understand the first condition, note that we are aiming to obtain a 
deterministic expression for the limiting population density. 
It is helpful to think
about a classical Wright-Fisher model (with no spatial structure and just two types, say). 
We know then that if the
timescale $\theta$ is on the same order as population size $N$, we see stochastic
fluctuations in the frequencies of the two types in the limit as $N\to\infty$; to 
obtain a deterministic limit, we look over timescales that are short relative to population
size. In our setting, the total population size is replaced by the local population
size, as measured by convolution with $\rho_{\epsilon}$, which we expect to be 
of order $N\epsilon^d$, and so in order to ensure a deterministic limit we 
take $\theta/(N\epsilon^d)\to 0$.
Concretely,
since $\|\rho^\epsilon_F\| \propto \epsilon^{-d/2}$,
the quadratic variation of $\rho^\epsilon_F\!*\!\eta(x)$
is $\mathcal{O}(\theta/(N\epsilon^d))$.

@@

In this section we present formal proofs of the first two of our
three scaling limits.
%following sections we present formal proofs of our three scaling limits.
In Subsection~\ref{sec:population_density_proof} 
we prove Theorem~\ref{thm:nonlocal_convergence}, to obtain (both
stochastic and deterministic) limits in
which interactions between individuals in the population are nonlocal.
In Subsection~\ref{subsec:nonlocal to local}
we show how, in two important examples in which the nonlocal
limit is respectively a deterministic solution to a non-local equation of 
reaction-diffusion type and a deterministic solution to a nonlocal 
porous medium equation with an additional logistic growth term, 
one can pass to a further 
limit to obtain a classical PDE.
%In subsequent sections (Sections~\ref{subsec:one step convergence proof} and~\ref{sec:lookdown_proofs})
%we prove our third scaling limit, in which 
%we pass from our population model 
%directly to a reaction-diffusion equation without going via a nonlocal limit,
%and finally our results concerning both convergence of the lookdown representation
%and characterization of the lineage process in the deterministic case.

@@

\begin{proof}[Proof of Lemma~\ref{lem:conditions on r}:]
Here, we need to prove that $|\gamma(x, \eta) B^\theta_f(x, \eta)|$ is bounded,
uniformly over $x$ and $\eta$.
First suppose that
assumption~\ref{control through r} of  
Lemma~\ref{lem:conditions on r} is satisfied. \revpoint{1}{28}
We write
\begin{align*}
    r(y,\eta) f(y) - r(x,\eta) f(x)
&=
    r(y,\eta) (f(y) - f(x))
    + (r(y,\eta) - r(x,\eta)) f(x) 
\\ &=
    r(y,\eta) \left(
        \sum_i (y-x)_i \partial_{x_i} f(x)
        + \sum_{ij} (y-x)_i (y-x)_j \partial_{x_i x_j} f(z_1)
    \right)
\\ &\qquad {}
    + f(x) \left(
        \sum_i (y-x)_i \partial_{x_i} r(x,\eta)
        + \sum_{ij} (y-x)_i (y-x)_j \partial_{x_i x_j} r(z_2,\eta)
    \right)
 \\ &=
    \left( r(x,\eta) + \sum_j (y-x)_j \partial_{x_j} r(z_3,\eta) \right)
    \left(
        \sum_i (y-x)_i \partial_{x_i} f(x)
    \right)
\\ &\qquad {}
    + r(y,\eta) 
        \sum_{ij} (y-x)_i (y-x)_j \partial_{x_i x_j} f(z_1)
\\ &\qquad {}
    + f(x) \left(
        \sum_i (y-x)_i \partial_{x_i} r(x,\eta)
        + \sum_{ij} (y-x)_i (y-x)_j \partial_{x_i x_j} r(z_2,\eta)
    \right) ,
\end{align*}
for some $z_i = \kappa_i x + (1-\kappa_i) y$.
Integrating this against $q(x, dy)$, we get that
\begin{align*}
& \bigg| \theta \int \left(
    r(y,\eta) f(y) - r(x,\eta) f(x)
\right) q_\theta(x, dy) \bigg|
\\ &\qquad \le
    \bigg| \sum_i \left( r(x,\eta) \partial_{x_i} f(x) + f(x) \partial_{x_i} r(x,\eta) \right)
    \theta \int (y-x)_i q_\theta(x, dy) \bigg|
\\ &\qquad \qquad {} +
    \bigg| f(x) 
    \theta \int \sum_{ij} \partial_{x_i x_j} r(z_2,\eta) (y-x)_i (y-x)_j q_\theta(x, dy) \bigg|
\\ &\qquad \qquad {} +
    \bigg| \theta \int \sum_{ij} (y-x)_i (y-x)_j 
        \left(
            \partial_{x_i} f(x) \partial_{x_j} r(z_3,\eta)
            + r(y,\eta) \partial_{x_ix_j} f(z_1)
        \right) q_\theta(x, dy) \bigg|
\end{align*}
Since $q_\theta(x,dy)$ is the density of a Gaussian with mean $\meanq(x)/\theta$
and covariance $\covq(x)/\theta$, and both $\meanq(x)$ and $\covq(x)$ are uniformly bounded,
so that $\theta \int (y-x)_i q_\theta(x,dy)$ is bounded as well.
Furthermore, a change of variables that diagonalizes $\covq(x)$
shows for any $g : \IR^d \to \IR^{d + d}$,
that if $C_g = \sup_y \sup_{\|z\| = 1} \sum_{ij} g(y)_{ij} z_i z_j$
and $\lambda_* = \sup_y \sup_{\|z\|=1} \sum_{ij} \covq(y)_{ij} z_i z_j$
then
$$ \theta \int \sum_{ij} g(y)_{ij} (y-x)_i (y-x)_j q_\theta(x, dy) \le C_g \lambda_* .  $$
Condition~\ref{control through r} 
gives uniform bounds on the derivatives of $r(x,\eta) = r(x,\rho_r*\eta(x))$
in this expression and so, provided $f$ also has uniformly bounded first and second derivatives,
we have a bound of the form
\begin{align*}
    |B^\theta_f| \leq K_1 + K_2 |f(x)| ,
\end{align*}
for suitable constants $K_1$, $K_2$ that depend only on the derivatives of $f$.

@@

Similarly,
\begin{align*}
    &
 \big|   \langle M^{N}(f) \rangle_{\tau + t} 
        - \langle M^{N}(f) \rangle_\tau \big|
    \\ & \qquad =
  \Big|  \int_{\tau}^{\tau + t}
        \frac{\theta}{N}
        \int_{\IR^d}
        \left\{
            \gamma(x, \eta^N_s)
            \int_{\IR^d} f^2(y) r(y, \eta^N_s) q_\theta(x, dy)
            +
            \mu_\theta(x, \eta^N_s) f^2(x)
        \right\}
        \eta^N_s(dx)
    ds \Big|
    \\ & \qquad =
   \Big| \int_{\tau}^{\tau + t}
        \frac{\theta}{N}
        \int_{\IR^d}
        \left\{
            \gamma(x, \eta^N_s)
            \left(
                2 f^2(x) r(x, \eta^N_s)
                +
                B^\theta_{f^2}(x, \eta^N_s)
            \right)
            -
            f^2(x) \frac{F(x, \eta^N_s)}{ \theta}
        \right\}
        \eta^N_s(dx)
    ds\Big| ,
\end{align*}
and so using the fact that $\theta/N \to \alpha < \infty$,
an argument entirely analogous to 
that for \eqref{eqn:eta_projections_goal1}
yields a $\delta_2$ for which \eqref{eqn:eta_projections_goal2} is
satsified. Taking $\delta=\min\{\delta_1,\delta_2\}$, the result follows.

>>>>

\KEYWORDS{population model; interacting superprocess; lookdown construction; porous medium equation; reaction-diffusion equation; travelling waves; genealogies; Fisher-KPP equation}
\AMSSUBJ{60J25 ; 92D10 ; 92D15 ; 92D25 ; 92D40}
\AMSSUBJSECONDARY{60F05 ; 60G09 ; 60G55 ; 60G57 ; 60H15 ; 60J68}

@@

The history of a natural population is often only accessible indirectly, 
through patterns of genetic diversity:
from genetic data, one can try to infer the genealogical trees that describe how
individuals in a sample from the population are related, and these have been shaped 
by its history (see e.g., \citep{neigel1993application,kelleher2019inferring}).
It is therefore of interest to establish information about the distribution of genealogical trees 
under our population model, which we do with a lookdown construction. 
Lookdown constructions were first introduced in~\cite{donnelly/kurtz:1996} to provide 
a mechanism for retaining information about genealogical relationships between individuals
sampled from a population evolving according to the Moran model when passing to the 
infinite population limit.
Since then, they have been extended to a wide range of models. Of particular
relevance to our work here are the papers ~\cite{kurtz/rodrigues:2011} 
and~\cite{etheridge/kurtz:2019}, in which
lookdown constructions are provided for a wide variety of population models, including
spatially structured branching processes.

@@

In what follows we first study various scaling limits of the spatial population process,
and then turn our attention to lineages traced back through these populations.
First, in Section~\ref{sec: Model and main results},
we describe the model and the main results,
Theorems~\ref{thm:nonlocal_convergence}, \ref{thm:local_convergence}, and~\ref{thm:lineages}.
Next, in Section~\ref{sec:applications}, we discuss a few striking consequences of these results
regarding the behavior of genealogies in travelling waves,
the appearance of periodic ``clumps'' in seemingly homogeneous population models,
and identifiability of the underlying dynamics from a stationary population profile.
In Section~\ref{sec:heuristics}, we provide heuristic explanations
of why the theorems ought to be true,
and some key ideas behind them,
and in Section~\ref{sec:lookdown} we define and discuss the lookdown construction.
Proofs of the results begin in Section~\ref{sec:proofs},
which proves results for population models with nonlocal interactions,
while Section~\ref{subsec:one step convergence proof} gives the more difficult proof
for the case when interaction distances also go to zero in the limit.
Finally, Section~\ref{sec:lookdown_proofs} gives proofs for convergence of the lookdown process
and the associated results for the motion of lineages.
The Appendix contains a few more technical and less central lemmas.
The results are illustrated in a few places with individual-based simulations,
made using SLiM \citep{haller_slim_2019},
but these are provided for visualization and we do not embark on numerical study.

@@

\paragraph{Scaling limits with nonlocal interactions:}
Recall that the process $(\eta_t^N)_{t\geq 0}$ takes its values in the space
${\mathcal D}_{[0,\infty)}(\measures)$ of c\`adl\`ag
paths on $\measures$. We endow $\measures$ with the topology of weak convergence
and 
${\mathcal D}_{[0,\infty)}(\measures)$ with the Skorohod topology.
A sequence of processes taking values in 
${\mathcal D}_{[0,\infty)}(\measures)$ is said to be tight if the corresponding
sequence of distributions is tight,
i.e., if any infinite subsequence has a weakly convergent subsubsequence.
Our first main result establishes tightness of our rescaled population processes in 
the case in which interactions remain nonlocal under the scaling, and 
characterises limit points as solutions to a martingale problem.

@@

Now that we have established what we can say about how 
population density changes with time,
we turn to results on ancestral lineages,
i.e., how genealogical ancestry can be traced back across the landscape.
Informally,
a \emph{lineage} $(L_t^N)_{t \ge 0}$,
begun at a spatial location $L_0^N = x$
where there is a focal individual in the present day,
can be obtained for each time $t$ by
setting $L_t^N$ to be the spatial location of the individual alive at time $t$
before the present from
whom the focal individual is descended.
Since in our model individuals have only one parent, this is unambiguous.
Although we did not explicitly retain such information,
it is clear that for finite $N$, since individuals are born one at a time, 
one could construct the lineage $(L_t^N)_{t=0}^T$
given the history of the population $(\eta^N_t)_{t = 0}^T$,
for each starting location to which $\eta^N_T$ assigns positive mass.
It is less clear, however, how to rigorously retain such information when we 
pass to the scaling limit.

However, the \emph{lookdown construction} in Section~\ref{sec:lookdown}
does just this -- \llabel{whatis_lineage}
the construction enables us to recover information about ancestry
in the infinite population limit,
and thus gives a concrete meaning to $(L_t)_{t \ge 0}$. \llabel{Linf_mention}
Roughly speaking,
each particle is assigned a unique ``level'' from $[0,\infty)$
that functions as a label and thus allows reconstruction of lineages.
The key to the approach is that levels are assigned in such a way as to be exchangeable, 
so that sampling a finite number, $k$ say, of
individuals from a given 
region is equivalent to looking at the individuals in that region with the $k$ lowest levels.
Moreover, as we pass to the infinite population limit,
the collection of (individual, level) pairs converges,
as we show in Theorem~\ref{thm:lookdown_convergence}.
See~\citet{etheridge/kurtz:2019} for an introduction to these ideas.
In particular, even in the infinite
population limit, we can sample an individual
from a region (it will be the individual in that region with the lowest level) 
and trace its line of descent. 
This will allow us to calculate, for each $x$ and $y\in\IR^d$, the proportion of
the population at location $x$ in the present day population that is descended from 
a parent who was at location $y$ at time $t$ in the past. To make
sense of this in our framework, in Section~\ref{sec:lineages_proof}, 
we justify a weak reformulation of this idea.

@@

Ancestry in travelling waves for populations described by reaction-diffusion equations
has been studied before by various authors, \revpoint{1}{16}
although most work assumes the diffusion term is linear
(in our notation, $r$ and $\gamma$ are constant,
but $F$ can depend on density).
For instance, following \citet{hallatschek/nelson:2008},
various authors (e.g., \citet{roques2012allee,birzu/hallatschek/korolev:2017})
describe genetic diversity
and ancestry in (possibly noisy) travelling waves for reaction-diffusion equations
in a situation that covers our first two examples below.
However, we are not aware of previous work covering the third case 
with nonlinear diffusive term (the porous medium equation).

@@

% % Removing this, since it's not actually that interesting:
% It is interesting to compare the stationary distribution we have obtained here
% to the expression that we'd get by setting $\meanq = -\wavespeed$
% and using Corollary~\ref{cor:stationary_dist},
% i.e., by giving each offspring a mean displacement that offsets the motion of the wave
% (The model is different because this motion only affects offspring, not adults.)
% In the Fisher-KPP and Allen-Cahn cases above, we get the same expressions,
% but this is only the case because $r\equiv\gamma\equiv1$ in both.
% For the PME with $\meanq = 1$ we have $\DG^* = \Delta + \grad$ 
% and so the equation solved by the population density is
% $$
%     \partial_t \varphi
%     =
%     \Delta (\varphi^2) + \grad (\varphi^2) + \varphi (1 - \varphi) ,
% $$
% which has a travelling wave solution of the same shape but moving more slowly,
% $\varphi(x,t) = w^{P}(a(x-2at))$, with $a=(\sqrt{5}-1)/4$,
% and a stationary distribution of the lineage relative to the wavefront of
% (NOT RIGHT:) $\pi(\xi) \propto e^{3\xi/2} (1 - e^{\xi/2})^3$.
% 
% \comment{Details to back that last statement up, which we probably want to comment out?}
% (We can see that $w^P(x-t/2)$ is a travelling wave solution by using that
% $\partial_x w^P = -(1 - w^P)/2$;
% and then the generator of a linage in the moving frame is
% $\Lgen_{P'} = \varphi \partial_x^2 + (4 \partial_x \varphi + 1/2 + \varphi) \partial_x$,
% which has speed measure proportional to (NOT RIGHT)
% $\exp(\int^x 3(1/2 dx + d \log \varphi)) / \varphi$.)
% \begin{align*}
% m_{P'}(\xi)
%     &\propto \frac{1}{2(1-e^{\xi/2})} \exp\left(
%         \frac{3}{2} \int^\xi \left\{
%             1 - \frac{e^{x/2}}{1 - e^{x/2}}
%         \right\}dx
%     \right) .
% \end{align*}
% \comment{Alison says:
%  If one tries a solution of the form $(1-\exp(a(x-ct)))_+$, then we find speed $c=2a$ and $a$ solves $4a^2+2a -1=0$ (I think - at least I got that twice). 
% }
% \comment{
% Checking that:
% $$(2ff')' + 2 b f f' = 2 f'^2 + 2 ff'' + 2 b f f'$$
% and so if $f' = -a(1-f)$ and thus $f'' = -a^2(1-f)$ then this is
% \begin{align*}
%     (f^2)'' + b (f^2)' + f(1-f)
%     &= 2 (a^2(1-f)^2 - a^2f(1-f) - abf(1-f)) + f(1-f) \\
%     &= (1-f) (2a^2(1-f) + (-2a^2-2ab+1)f) \\
%     &= (1-f) (2a^2 + (-4a^2-2ab+1)f)
% \end{align*}
% Now, solutions to $4a^2+2ab-1=0$ are $a_{\pm}=(-b \pm \sqrt{b^2+4})/4$ and for these
% \begin{align*}
%     c_\pm 
%     &= 2a_\pm \\
%     &= \frac{-b \pm \sqrt{b^2+4}}{2} .
% \end{align*}
% }

@@

It is natural for applications to wonder about identifiability:
when can the observed quantities like population density
or certain summaries of lineage movement uniquely determine
the underlying demographic parameters?
Consider a deterministic,
continuous population generated by parameters $\gamma$, $r$, and $F$,
with $\meanq = 0$ and $\covq = 2I$.
Suppose it has a stationary profile $w(x)$, that must satisfy
$$
   r \Delta(\gamma w) + F w = 0 .
$$
It is easy to see that $w$ does not uniquely specify $\gamma$, $F$, and $r$:
let $\lambda(x)$ be a smooth, nonnegative function on $\IR^d$,
and let $\widetilde{r}(x, m) = \lambda(x) r(x, m)$ and $\widetilde{F}(x, m) = \lambda(x) F(x, m)$
(and, let $\widetilde{\gamma} = \gamma$).
Since $\mu = r \gamma - F/\theta$,
this corresponds to multiplying both establishment probabilities and death rates by $\lambda$.
Then the population with parameters $\widetilde{\gamma}$, $\widetilde{r}$, and $\widetilde{F}$
has the same stationary profile(s) as the original population.

@@

In this section we perform some preliminary calculations and use them 
to provide heuristic arguments for our main results,
to build intuition before the proofs.

@@

\paragraph{One-step convergence:} \label{sec:one_step_heuristics} \llabel{one_step_heuristics}
In order to pass directly to a classical PDE limit
in Theorem~\ref{thm:local_convergence}
we impose the stronger condition
that $\theta/(N\epsilon^d) \to 0$ and also require that
$\theta\epsilon^2\to\infty$. 
Recall that in this case, we take $\rho_F^\epsilon$ 
to be a symmetric Gaussian density with variance 
$\epsilon^2$. The condition $\theta\epsilon^2\to\infty$ 
ensures that $\epsilon^2$ is large enough relative to $1/\theta$
that the regularity gained by smoothing our population density by convolution with
$\rho_\epsilon$ is preserved under the dynamics dictated by $q_{\theta}$.
To understand the first condition, note that we are aiming to obtain a 
deterministic expression for the limiting population density. 
It is helpful to think
about a classical Wright-Fisher model (with no spatial structure and just two types, say). 
We know then that if the
timescale $\theta$ is on the same order as population size $N$, we see stochastic
fluctuations in the frequencies of the two types in the limit as $N\to\infty$; to 
obtain a deterministic limit, we look over timescales that are short relative to population
size. In our setting, the total population size is replaced by the local population
size, as measured by convolution with $\rho_{\epsilon}$, which we expect to be 
of order $N\epsilon^d$, and so in order to ensure a deterministic limit we 
take $\theta/(N\epsilon^d)\to 0$.
% Concretely,
% the quadratic variation of $\rho^\epsilon_F\!*\!\eta(x)$
% is $\mathcal{O}(\theta/(N\epsilon^d))$.

@@

In this section we present proofs of the first two of our
three scaling limits.
In Subsection~\ref{sec:population_density_proof} 
we prove Theorem~\ref{thm:nonlocal_convergence}, to obtain (both
stochastic and deterministic) limits in
which interactions between individuals in the population are nonlocal.
In Subsection~\ref{subsec:nonlocal to local}
we show how, in two important examples in which the nonlocal
limit is respectively a deterministic solution to a non-local equation of 
reaction-diffusion type and a deterministic solution to a nonlocal 
porous medium equation with an additional logistic growth term, 
one can pass to a further 
limit to obtain a classical PDE.
%In subsequent sections (Sections~\ref{subsec:one step convergence proof} and~\ref{sec:lookdown_proofs})
%we prove our third scaling limit, in which 
%we pass from our population model 
%directly to a reaction-diffusion equation without going via a nonlocal limit,
%and finally our results concerning both convergence of the lookdown representation
%and characterization of the lineage process in the deterministic case.

@@

\begin{proof}[Proof of Lemma~\ref{lem:conditions on r}:]
Here, we need to prove that $|\gamma(x, \eta) B^\theta_f(x, \eta)|$ is bounded,
uniformly over $x$ and $\eta$.
Note that Conditions~\ref{def:model_setup} assume nothing about $\eta$,
and so, for instance, although $r(x,m)$ has uniformly bounded derivatives,
it might still be the case that $r(x,\eta) = r(x, \smooth{r}\eta(x))$ changes arbitrarily rapidly;
the additional conditions of the Lemma prevent this from happening.

First suppose that
assumption~\ref{control through r} of  
Lemma~\ref{lem:conditions on r} is satisfied. \revpoint{1}{28}
We write
\begin{align*}
    r(y,\eta) f(y) - r(x,\eta) f(x)
&=
    r(y,\eta) (f(y) - f(x))
    + (r(y,\eta) - r(x,\eta)) f(x) 
\\ &=
    r(y,\eta) \left(
        \sum_i (y-x)_i \partial_{x_i} f(x)
        + \sum_{ij} (y-x)_i (y-x)_j \partial_{x_i x_j} f(z_1)
    \right)
\\ &\qquad {}
    + f(x) \left(
        \sum_i (y-x)_i \partial_{x_i} r(x,\eta)
        + \sum_{ij} (y-x)_i (y-x)_j \partial_{x_i x_j} r(z_2,\eta)
    \right)
 \\ &=
    \left( r(x,\eta) + \sum_j (y-x)_j \partial_{x_j} r(z_3,\eta) \right)
    \left(
        \sum_i (y-x)_i \partial_{x_i} f(x)
    \right)
\\ &\qquad {}
    + r(y,\eta) 
        \sum_{ij} (y-x)_i (y-x)_j \partial_{x_i x_j} f(z_1)
\\ &\qquad {}
    + f(x) \left(
        \sum_i (y-x)_i \partial_{x_i} r(x,\eta)
        + \sum_{ij} (y-x)_i (y-x)_j \partial_{x_i x_j} r(z_2,\eta)
    \right) ,
\end{align*}
for some $z_i = \kappa_i x + (1-\kappa_i) y$.
Integrating this against $q(x, dy)$, we get that
\begin{align*}
& \bigg| \theta \int \left(
    r(y,\eta) f(y) - r(x,\eta) f(x)
\right) q_\theta(x, dy) \bigg|
\\ &\qquad \le
    \bigg| \sum_i \left( r(x,\eta) \partial_{x_i} f(x) + f(x) \partial_{x_i} r(x,\eta) \right)
    \theta \int (y-x)_i q_\theta(x, dy) \bigg|
\\ &\qquad \qquad {} +
    \bigg| f(x) 
    \theta \int \sum_{ij} \partial_{x_i x_j} r(z_2,\eta) (y-x)_i (y-x)_j q_\theta(x, dy) \bigg|
\\ &\qquad \qquad {} +
    \bigg| \theta \int \sum_{ij} (y-x)_i (y-x)_j 
        \left(
            \partial_{x_i} f(x) \partial_{x_j} r(z_3,\eta)
            + r(y,\eta) \partial_{x_ix_j} f(z_1)
        \right) q_\theta(x, dy) \bigg| .
\end{align*}
Since $q_\theta(x,dy)$ is the density of a Gaussian with mean $\meanq(x)/\theta$
and covariance $\covq(x)/\theta$, and both $\meanq(x)$ and $\covq(x)$ are uniformly bounded,
$\theta \int (y-x)_i q_\theta(x,dy)$ is bounded as well.
Furthermore, a change of variables that diagonalizes $\covq(x)$
shows for any $g : \IR^d \to \IR^{d + d}$,
that if $C_g = \sup_y \sup_{\|z\| = 1} \sum_{ij} g(y)_{ij} z_i z_j$
and $\lambda_* = \sup_y \sup_{\|z\|=1} \sum_{ij} \covq(y)_{ij} z_i z_j$
then
$$ \theta \int \sum_{ij} g(y)_{ij} (y-x)_i (y-x)_j q_\theta(x, dy) \le C_g \lambda_* .  $$
Condition~\ref{control through r} 
gives uniform bounds on the derivatives of $r(x,\eta) = r(x,\rho_r*\eta(x))$
in this expression and so, provided $f$ also has uniformly bounded first and second derivatives,
we have a bound of the form
\begin{align*}
    |B^\theta_f| \leq K_1 + K_2 |f(x)| ,
\end{align*}
for suitable constants $K_1$, $K_2$ that depend only on the derivatives of $f$.

@@

Similarly,
\begin{align*}
    &
 \big|   \langle M^{N}(f) \rangle_{\tau_N + t} 
        - \langle M^{N}(f) \rangle_{\tau_N} \big|
    \\ & \qquad =
  \Big|  \int_{\tau_N}^{\tau_N + t}
        \frac{\theta}{N}
        \int_{\IR^d}
        \left\{
            \gamma(x, \eta^N_s)
            \int_{\IR^d} f^2(y) r(y, \eta^N_s) q_\theta(x, dy)
            +
            \mu_\theta(x, \eta^N_s) f^2(x)
        \right\}
        \eta^N_s(dx)
    ds \Big|
    \\ & \qquad =
   \Big| \int_{\tau_N}^{\tau_N + t}
        \frac{\theta}{N}
        \int_{\IR^d}
        \left\{
            \gamma(x, \eta^N_s)
            \left(
                2 f^2(x) r(x, \eta^N_s)
                +
                B^\theta_{f^2}(x, \eta^N_s)
            \right)
            -
            f^2(x) \frac{F(x, \eta^N_s)}{ \theta}
        \right\}
        \eta^N_s(dx)
    ds\Big| ,
\end{align*}
and so using the fact that $\theta/N \to \alpha < \infty$,
an argument entirely analogous to 
that for \eqref{eqn:eta_projections_goal1}
yields a $\delta_2$ for which \eqref{eqn:eta_projections_goal2} is
satsified. Taking $\delta=\min\{\delta_1,\delta_2\}$, the result follows.
